---
title: "CSCI 385 - Introduction to Data Science"
subtitle: "Introduction to R"
author: "Derek Borders"
date: "Last compiled on `r format(Sys.time(), '%m/%d/%y')`"
output: 
    html_document:
        theme: sandstone        # Reduces eye strain
        #toc: true               # Table of contents
        #toc_float: true         # ^ in a sidebar
        code_folding: show      # Collapsable code chunks
        code_download: true     # Page contains code for reader to examine

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)          # default
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment=NA)
options(digits = 2)

library(Rcpp)                        # for Cpp
```




# {.tabset .tabset-fade .tabset-pills} 

Solve the given problems from scratch and test each solution with 3-5 simple test cases showing that the solution behaves as expected. Target important edge cases.

##  {}

### Sum of Multiples  

*(5 pts) Write a function sum_mults(nums, n) that returns the sum of all multiples of values in the vector nums less than n. For example, sum_mults(c(3,5), 30) should return 195. Assume that the elements of nums are positive integers.*

```{r}

```


##  {}

### Collatz Conjecture 


```{r}

```



##  {}

### Reverse


```{r}

```



##  {}

### Drop


```{r}

```



##  {}

### Intersect


```{r}

```



##  {}

### Filter


```{r}

```



##  {}

### Fibonacci


```{r}

```



##  {}

### Shift Right


```{r}

```



##  {}

### Remove Consecutive Dupes


```{r}

```



##  {}

### Even Fibonacci


```{r}

```


