---
title: "CSCI 385 - Introduction to Data Science"
subtitle: "Introduction to R"
author: "Derek Borders"
date: "Last compiled on `r format(Sys.time(), '%m/%d/%y')`"
output: 
    html_document:
        theme: sandstone        # Reduces eye strain
        #toc: true               # Table of contents
        #toc_float: true         # ^ in a sidebar
        code_folding: show      # Collapsable code chunks
        code_download: true     # Page contains code for reader to examine

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # default
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment = NA)
options(digits = 2)

library(Rcpp) # for Cpp
```




# {.tabset .tabset-fade .tabset-pills} 

Solve the given problems from scratch and test each solution with 3-5 simple test cases showing that the solution behaves as expected. Target important edge cases.

## 1 {}

### Sum of Multiples  

(5 pts) Write a function `sum_mults(nums, n)` that returns the sum of all multiples of values in the vector `nums` less than `n`. For example, `sum_mults(c(3,5), 30)` should return `195`. Assume that the elements of `nums` are positive. 

### Solution

The naive solution here will be to use nested loops to iterate over a sequence from 1 to n and add any value which is a multiple of any value in num.

```{r}
# Naive Solution (nested loop)

sum_mults <- function(nums, n) {
    if (n<1){return(0)}
    sum <- 0
    for (i in 1:n-1){
        for (num in nums){
            if (i %% num == 0){
                sum <- sum + i
                break
            }
        }
    }
    return(sum)
}
```  

### Testing

For test cases, we will first test the given example: `sum_mults(c(3,5), 30)`. This is a good test because one possible failure condition would be adding multiples of more than one factor more than once (15 would be counted twice for 210).

```{r}
sum_mults(c(3,5), 30) # Expect 195
```


Next, we will test a case with no multiples by choosing `nums` which are greater than the chosen limit. 

```{r}
sum_mults(c(101,102), 30) # Expect 0
```

Next, we will test a vector with duplicate values, though this shouldn't raise any new failure conditions, but we want to be sure duplicates don't break it in an unforseen way. 

```{r}
sum_mults(c(3, 3, 5), 31) # Expect 225
```

Next, we will check what happens when we pass in an empty vector. 

```{r}
sum_mults(c(NULL), 30) # Expect 0
```

Next, we will test passing a limit that is less than 1.

```{r}
sum_mults(c(3,5), -30) # Expect 0
```



## 2 {}

### Collatz Conjecture 
(5 pts) Given any positive integer n, define  

\[ 
    f(n) :=
    \begin{cases} 
      n/2, & \verb|if |n \verb| is even| \\
      3n+1, & \verb|if |n \verb| is odd|
   \end{cases}
\]

The [Collatz conjecture](https://en.wikipedia.org/wiki/Collatz_conjecture) states that the sequence 

\[ 
    a_i :=
    \begin{cases} 
      n, & \verb|if |i=1\\
      f(a_\verb|i-1|), & \verb|if |i>1
   \end{cases}
\]

eventually reaches 1. Write a function `collatz_len(n)` that determines the first $i$ for which $a_i=1$ for a given $n$. 

For example, when $n=17$, the sequence $a_i$ begins $17, 52, 26, 13, 40, 20, 10, 5, 16, 8, 4, 2, 1, ...$ and $a_13 = 1$. Thus, `collatz_len(17)` should return ~~13~~ 12.

### Solution

First, we will write the function itself, and call it `next.num()`. 

```{r}
# The inner function
next.num <- function(n){
    if (n %% 2){
        return(3*n+1)
    }else{
        return(n/2)
    }
}
```

We can quickly test `next.num()` with 4 and 3 to check that it works for odd and even numbers.  
```{r}
next.num(4) # Expect 2
next.num(3) # Expect 10
```  

Next, we build the function to determine a number's Collatz Length (or height).

```{r}
collatz_len <- function(n){
    if (n<1){return(-1)}
    num <- n
    c.len <- 0
    while (num > 1){
        num <- next.num(num)
        c.len <- c.len + 1
    }
    return(c.len)
}

# collatz_len(943)
```

### Testing

For testing, we will again start with the given example. For the Collatz Length of 17, we expect ~~13~~ 12 and we get: `r collatz_len(17)`

Next, we will test a number with a known Collatz Height from a table in an interesting [paper](https://www.sciencedirect.com/science/article/pii/0012365X9390240T?via%3Dihub) I came across that explores the interesting phenomenon of stretches of numbers with identical Collatz Heights. The first such stretch listed is the seven numbers starting with 943 which all have a height of 36.

Running our implementation on 943, we get a height of `r collatz_len(943)` as expected.

Next we will test for known invalid inputs, 0 or a negative number. We have written this implementation to return -1 for these cases so we expect this result. For 0, we get `r collatz_len(0)`, for -10 we get `r collatz_len(-10)`.

Finally, we will test the first 72 positive integers against a known list found [here](https://oeis.org/A006577/list).

```{r}
# Known list
known72 <- c(0,1,7,2,5,8,16,3,19,6,14,9,9,17,17,4,12,20,20,7, 7,15,15,10,23,10,111,18,18,18,106,5,26,13,13,21, 21,21,34,8,109,8,29,16,16,16,104,11,24,24,24,11, 11,112,112,19,32,19,32,19,19,107,107,6,27,27,27, 14,14,14,102,22)

# Test list
test72 <- rep(0, 72)                # Initialize vector of 0s
for (i in 1:72){                    # Loop through integers to 72
    test72[i] <- collatz_len(i)     # Store height for each integer
}    

# Compare lists
success <- identical(test72, known72)
success
```


## 3 {}

### Reverse  
(5 pts) Write a function `reverse(v)` that reverses the vector `v`. So `reverse(c(1,2,3))` should return `c(3,2,1)`. `reverse` should return `NULL`, not `NA`, when `v` is `c()`.

```{r}
reverse <- function(v){
    len <- length(v)
    if (len < 2){return(v)}
    for(i in 1:len/2){
        start <- i
        end <- (len +1)-1
        temp <- v[i]
        v[i] <- v[len-i+1]
        v[len-i+1] <- temp
    }
}
reverse(c())
reverse(c('palindrome'))
reverse(c(1,2))
```



## 4 {}

### Drop


```{r}

```



## 5 {}

### Intersect


```{r}

```



## 6 {}

### Filter


```{r}

```



## 7 {}

### Fibonacci


```{r}

```



## 8 {}

### Shift Right


```{r}

```



## 9 {}

### Remove Consecutive Dupes


```{r}

```



## 10 {}

### Even Fibonacci


```{r}

```


